\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-08-07 coqimpl.tex
Save. Changing family compilation part.
Rewrite CoqImpl to coincide with the latest implementation.

\textbf{Family compiled into module.} The main idea of our plugin is to support family by compiling
family components into Coq modules, family ``types'' into module types, and the context of a given term
into parameters of the module. We need to faithfully
reflect the \textit{late binding} nature of families: for each defining
field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
\cL \sigma. T$. However, instead of using Coq's universal quantifier, we
choose module and module parameters to achieve this wrapping---each
family type in the context will be one module parameter
with the compiled module type, and thus each field of the family will
compile into a functor parametrized by its context. Doing so we can also
get rapid feedback from type checking when the users are defining each
field interactively with Coq.
% Insert one pseudo-code example for explanation of the mechanism
\begin{figure}[!htb]
  \begin{minipage}[t]{0.30\linewidth}
\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
Family A.
Field a : nat := 1.
Field b : a = a := eq_refl.
Final Field a' : nat := 1. 
EndFamily.
\end{lstlisting}
  \end{minipage}%\YZ{Can the self's be omitted?}\EDJreply{Yes. But in our current plugin implementation it is not omitted. I think altering example is fine. If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}\EDJreply{Omitting self_ is easy and not much engineering effort is required. But omitting self_ will lead to `A.a' instead of `a' actually. So if we really want `a' as you last time mentioned, then a resolution is required and more engineering effort is required. But I don't see any technical difficulties. I have removed all the self, please check.}
  \begin{minipage}[t]{0.35\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module a_4 (self_A: EmptySig).
Definition a : nat := 1. End a_4.
Module Type a_5 (self_A: EmptySig).
Parameter (a : nat). End a_5.
Module Type a_6 := a_5 EmptyMod.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.3\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module b_7 (self_A: a_6).
Definition b 
  : self_A.a = self_A.a 
  := eq_refl. End b_7.

@@

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: immediately after the user inputs the line "Field a : nat := ..", our plugin will
translate this statement into the "Module a_4", "Module Type a_5, a_6"
three components.
{The "EmptySig" is just an empty module type.} Then,
these modules and module types will be generated and type-checked by Coq
immediately. The compiled "Module a_4" is then part of the context of
"Field b" (as the module type of "self_A") for "b" to refer to. The
internal representation of family "A" is the list of these
compiled modules and module types. We can achieve incremental
checking as in \ref{chg:software-engineering}; overriding is also directly supported---for example, if we
want to override "Field a", we simply replace that compiled module with
the new module inside the list. This compiled module type serves as an
abstraction that enables overriding.

This is achieved by a complete compilation from a family to a module,
via aggregating the list of the modules in the internal representation:
the plugin will repeatedly using Coq's "Include" command as the "Module
A" as \cref{fig:plugin-example1} illustrates.
References to family "A" outside "A" will be compiled to references to
this Coq module "A".

Note that, in our example, we include two
functors without arguments---this is due to a peculiarity of Coq's
"Include": uninstantiated parameters will be automatically
instantiated by the appropriate fields in the surrounding defining
module.
For example, "b_7" requires a module of type "a_6". Inside module "A",
when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat}
inside the current surrounding "A". Then when we
\mintinline{Coq}{Include b_7} without specifying the module parameter,
Coq will try to find fields in the surrounding module (i.e., module
"A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by
the \mintinline{Coq}{a : nat} that is included earlier.
The resulting compiled module is just like mundane Coq's module and 
thus maintains all the computational mechanism from Coq, fulfilling
requirements in \ref{chg:software-engineering}.\YZ{
  Do I understand correctly that an alternative to individually
  compiling family members to modules is to compile a family as a whole,
  but this alternative would not satisfy the "incremental type checking"
  or "instant feedback" requirement?
}\YZreply{
  Looks like individually compiling family members addresses not only
  this requirement, but also the "self vs. consistency" challenge.
}\EDJreply{Yes. Indivially compiling family member can achieve both requirements}

@@

\textbf{Implementing "Final" and "Sealed Family".}
We implement "Final" by
exposing the whole definition directly into the compiled module type.
Doing so will prohibit overriding. Say, "bop", declared "Final" in
\cref{fig:plugin-example3}, is not overridable.
We require the users to manual decorate field as `Final' so that the
plugin will proceed with this special treatment.
%\YZ{This para is confusing: which kind of equality is used, propositional or judgmental? Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on talking about meta-theory. Now everything is about judgemental so I don't have to emphasize judgemental. I wanted to make the example small but I guess I can add one `Final' statement into our example. Don't resolve it until I add it into Figure 2.}

% \textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.}

The implementation of sealed family is not so different from that of the normal family, but we need to deal with the compiled module type. The compiled module type is the aggregation of the type instead of the concrete definition of the fields.
For example, the compiled module type of
"MonotonicF" in \cref{fig:plugin-example3} will be "{f : nat -> nat; p : monotonic f}". Thus any future overriding family will always have these two fields "f, p" with the corresponding type.\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin. }

>>>>

\textbf{Family compiled into module.} The main idea of our plugin is to support family by compiling
family components into Coq modules, family ``types'' into module types, and the context of a given term
into parameters of the module.

Since by default every field is inheritable, for each defining field (informally) $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
\cL \sigma. T$, so that field $t$ can be inherited to all the future family with strengthen context $\sigma^+$ with the help of weakening from $\sigma^+$ to $\sigma$.  However, instead of using Coq's universal quantifier, we choose module and module parameters to achieve this wrapping---each
family type in the context will be one module parameter
with the compiled module type, and thus each field of the family will
compile into a functor parametrized by its context. Doing so we can also
get rapid feedback from type checking when the users are defining each
field interactively with Coq. Note that, when compiling the context during the  compilation of a default (non-overridable)  field, all the default (non-overridable) former fields will expose their definition in the context while overridable fields will only expose their types.
% Insert one pseudo-code example for explanation of the mechanism
\begin{figure}[!htb]
  \begin{minipage}[t]{0.30\linewidth}
\begin{lstlisting}[language=Coq, escapeinside={@}{@}]
Family A.
Field a : nat := 1.
Field b : a = a := eq_refl.
Final Field a' : nat := 1. 
EndFamily.
\end{lstlisting}
  \end{minipage}%\YZ{Can the self's be omitted?}\EDJreply{Yes. But in our current plugin implementation it is not omitted. I think altering example is fine. If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}\EDJreply{Omitting self_ is easy and not much engineering effort is required. But omitting self_ will lead to `A.a' instead of `a' actually. So if we really want `a' as you last time mentioned, then a resolution is required and more engineering effort is required. But I don't see any technical difficulties. I have removed all the self, please check.}
  \begin{minipage}[t]{0.35\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module a_4 (self_A: EmptySig).
Definition a : nat := 1. End a_4.
Module Type a_6 (self_A: EmptySig).
Include a_4. End a_5.

\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.3\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module b_7 (self_A: a_6).
Definition b 
  : self_A.a = 1
  := eq_refl. End b_7.

@@

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: immediately after the user inputs the line "Field a : nat := ..", our plugin will
translate this statement into the "Module a_4", "Module Type a_6"
two components.
{The "EmptySig" is just an empty module type.} Then,
these modules and module types will be generated and type-checked by Coq
immediately. The compiled "Module a_6" is then part of the context of
"Field b" (as the module type of "self_A") for "b" to refer to, so that "b" knows "a"'s concrete definition. The
internal representation of family "A" is the list of these
compiled modules and module types. We can achieve incremental
checking as in \ref{chg:software-engineering}. This makes all kinds of inheritance plausible---for example, if we
want to override "Field a", we simply replace that compiled module with
the new module inside the list. 
% This compiled module type serves as an abstraction that enables overriding.

After closing the top family, the user can access family "A" like it is a module. This is achieved by a complete compilation from a family to a module,
via aggregating the list of the modules in the internal representation:
the plugin will repeatedly using Coq's "Include" command as the "Module
A" as \cref{fig:plugin-example1} illustrates.
References to family "A" outside "A" will be compiled to references to
this Coq module "A".

Note that, in our example, we include two
functors without arguments---this is due to a peculiarity of Coq's
"Include": uninstantiated parameters will be automatically
instantiated by the appropriate fields in the surrounding defining
module.
For example, "b_7" requires a module of type "a_6". Inside module "A",
when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat := 1}
inside the current surrounding "A". Then when we
\mintinline{Coq}{Include b_7} without specifying the module parameter,
Coq will try to find fields in the surrounding module (i.e., module
"A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by
the \mintinline{Coq}{a : nat} that is included earlier.
The resulting compiled module is just like mundane Coq's module and 
thus maintains all the computational mechanism from Coq, fulfilling
requirements in \ref{chg:software-engineering}.\YZ{
  Do I understand correctly that an alternative to individually
  compiling family members to modules is to compile a family as a whole,
  but this alternative would not satisfy the "incremental type checking"
  or "instant feedback" requirement?
}\YZreply{
  Looks like individually compiling family members addresses not only
  this requirement, but also the "self vs. consistency" challenge.
}\EDJreply{Yes. Indivially compiling family member can achieve both requirements}

\textbf{Implementing "Overridable" and Overriding Chain.}
We implement "Overridable" by simply annotating the corresponding field type with a mark and its dependency. Then when compute the context during the compilation of the term for that field, we make sure only all the dependee and the non-overridable field will expose its definition in the context. Most parts of the implementation is quite similar to the default fields, except that when inheriting an overridable field, we need to make sure all its dependee is inherited earlier.

When aggregating the fields into a module, we will make sure all the overridable fields to expose its definition for simiplicity in the compilation.

@@

% \textbf{Implementing "Final" and "Sealed Family".}
% We implement "Final" by
% exposing the whole definition directly into the compiled module type.
% Doing so will prohibit overriding. Say, "bop", declared "Final" in
% \cref{fig:plugin-example3}, is not overridable.
% We require the users to manual decorate field as `Final' so that the
% plugin will proceed with this special treatment.
% %\YZ{This para is confusing: which kind of equality is used, propositional or judgmental? Where does 'Final' show up in the example?}\EDJreply{I rephrase it and remove the part on talking about meta-theory. Now everything is about judgemental so I don't have to emphasize judgemental. I wanted to make the example small but I guess I can add one `Final' statement into our example. Don't resolve it until I add it into Figure 2.}

% % \textbf{Special Fields: Overridable/Sealed Family Simulating Sigma Type.}

% The implementation of sealed family is not so different from that of the normal family, but we need to deal with the compiled module type. The compiled module type is the aggregation of the type instead of the concrete definition of the fields.
% For example, the compiled module type of
% "MonotonicF" in \cref{fig:plugin-example3} will be "{f : nat -> nat; p : monotonic f}". Thus any future overriding family will always have these two fields "f, p" with the corresponding type.\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin. }

% The implementation of sealed family is not so different from that of the normal family, but we need to deal with the compiled module type. The compiled module type is the aggregation of the type instead of the concrete definition of the fields.
% For example, the compiled module type of
% "MonotonicF" in \cref{fig:plugin-example3} will be "{f : nat -> nat; p : monotonic f}". Thus any future overriding family will always have these two fields "f, p" with the corresponding type.\YZ{Can an overriding, sealed family define new fields not in its parent?}\EDJreply{I don't find this a problem but I didn't support it in my plugin. }
